---
title: "d13C_STATS"
author: "Conner Mertz"
date: "2023-09-05"
output: html_document
---

This code tests for significant differences in measured d13C AA data. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# start by loading the necessary libraries
library(MASS)
library(stats)
library(car)
library(ellipse)
library(ggplot2)
library(ggnewscale)
library(tidyr)
library(ggplot2)

```


Read in your data
```{r}
# read in your data!
meta <- read.csv("/Users/connermertz/Dropbox (vlab)/Conner Mertz/UNM biology/Chapter 1 E1 vs E2/Methods/SIA/E1_E2_d13C Data/STATS/d13C_E1_E2_muscleCricketCornmeal_STATS_import.csv")

#make column names SampleID, Val13C, Leu13C..ect,and include any other meta data you want. For me that was Experiment # (E1 or E2), Percent_Protein (2.5P,5P or 10P), ect.. 

```

```{r}
# subsetting our data by creating separate dataframes

Muscle_Diet.ess <- subset(meta, select = c("SampleID","Type", "Experiment_Percent_Protein", "Treatment", "Percent_Protein", "Val13C","Leu13C", "Ile13C",	"Phe13C",	"Thr13C", "Lys13C"))


E1_.ess <- subset(Muscle_Diet.ess, Treatment == "Synthetic")
E2_.ess <- subset(Muscle_Diet.ess, Treatment == "Semi_Natural")


```


Lets do some statistical tests! Remember: 
 
-	 T-test (For 2 groups, compare means using t-tests (if data are Normally distributed, parametric test), or Mann-Whitney U test (For 2 groups, if data are not normally distributed, non parametric test) 
-	ANOVAs (for greater than 2 groups, normally distributed data, parametric test), or Kruskal-Wallis ANOVA+Wilcox Rank Sum test (for greater than 2 groups, not normally distributed data, non parametric test).

**Important Caveot** 
It is ok to use non- parametric tests on normally distributed data!  

##Test for differences in E1 casein vs E1 AAESS muscle values:
####################
1. Test for normality
```{r}
# test for normality (p-values above 0.05 indicate normality)
#"Val13C","Leu13C", "Ile13C",	"Phe13C",	"Thr13C", "Lys13C"
shapiro.test(E1_.ess$Val13C)
shapiro.test(E1_.ess$Leu13C)
shapiro.test(E1_.ess$Ile13C)
shapiro.test(E1_.ess$Phe13C)
shapiro.test(E1_.ess$Thr13C)
shapiro.test(E1_.ess$Lys13C)

```
None of these are normally distributed so we need to perform a -	Perform Kruskal-Wallis ANOVA+Wilcox Rank Sum test (for greater than 2 groups, not normally distributed data): 

#non-parametric tests for all AAESS
```{r}
#Kruskal-Wallis ANOVA+Wilcox Rank Sum test
Ile <- kruskal.test(E1_.ess$Ile13C ~ E1_.ess$Experiment_Percent_Protein, data = E1_.ess) #determines whether there is a difference in d13CIle in E1 among Percent Protein diets
Ile
#follow it up by pairwise comparison to see which groups are diff. 
pairwise.wilcox.test(E1_.ess$Ile13C, E1_.ess$Experiment_Percent_Protein,
                 p.adjust.method = "BH")

#Kruskal-Wallis ANOVA+Wilcox Rank Sum test
Leu <- kruskal.test(E1_.ess$Leu13C ~ E1_.ess$Experiment_Percent_Protein, data = E1_.ess) #determines whether there is a difference in d13CLeu in E1 among Percent Protein diets
Leu
#follow it up by pairwise comparison to see which groups are diff. 
pairwise.wilcox.test(E1_.ess$Leu13C, E1_.ess$Experiment_Percent_Protein,
                 p.adjust.method = "BH")

#Kruskal-Wallis ANOVA+Wilcox Rank Sum test
Val <- kruskal.test(E1_.ess$Val13C ~ E1_.ess$Experiment_Percent_Protein, data = E1_.ess) #determines whether there is a difference in d13CVal in E1 among Percent Protein diets
Val
#follow it up by pairwise comparison to see which groups are diff. 
pairwise.wilcox.test(E1_.ess$Val13C, E1_.ess$Experiment_Percent_Protein,
                 p.adjust.method = "BH")

#Kruskal-Wallis ANOVA+Wilcox Rank Sum test
Phe <- kruskal.test(E1_.ess$Phe13C ~ E1_.ess$Experiment_Percent_Protein, data = E1_.ess) #determines whether there is a difference in d13CPhe in E1 among Percent Protein diets
Phe
#follow it up by pairwise comparison to see which groups are diff. 
pairwise.wilcox.test(E1_.ess$Phe13C, E1_.ess$Experiment_Percent_Protein,
                 p.adjust.method = "BH")

#Kruskal-Wallis ANOVA+Wilcox Rank Sum test
Thr <- kruskal.test(E1_.ess$Thr13C ~ E1_.ess$Experiment_Percent_Protein, data = E1_.ess) #determines whether there is a difference in d13CThr in E1 among Percent Protein diets
Thr
#follow it up by pairwise comparison to see which groups are diff. 
pairwise.wilcox.test(E1_.ess$Thr13C, E1_.ess$Experiment_Percent_Protein,
                 p.adjust.method = "BH")

#Kruskal-Wallis ANOVA+Wilcox Rank Sum test
Lys <- kruskal.test(E1_.ess$Lys13C ~ E1_.ess$Experiment_Percent_Protein, data = E1_.ess) #determines whether there is a difference in d13CLys in E1 among Percent Protein diets
Lys
#follow it up by pairwise comparison to see which groups are diff. 
pairwise.wilcox.test(E1_.ess$Lys13C, E1_.ess$Experiment_Percent_Protein,
                 p.adjust.method = "BH")


```



###Now lets look for differences in just E2:
1.	Test for normality (Shapiro-Wilk normality test) d13C essential AA of muscle E2. (p value above 0.05 indicate normality):

```{r}
# test for normality (p-values above 0.05 indicate normality)
#"Val13C","Leu13C", "Ile13C",	"Phe13C",	"Thr13C", "Lys13C"
shapiro.test(E2_.ess$Val13C)
shapiro.test(E2_.ess$Leu13C)
shapiro.test(E2_.ess$Ile13C)
shapiro.test(E2_.ess$Phe13C)
shapiro.test(E2_.ess$Thr13C)
shapiro.test(E2_.ess$Lys13C)


```

Only a few of these are normally distributed so we need to Perform Kruskal-Wallis ANOVA+Wilcox Rank Sum test (for greater than 2 groups, not normally distributed data): 

#non-parametric tests for all AAESS
```{r}
#Kruskal-Wallis ANOVA+Wilcox Rank Sum test
Ile <- kruskal.test(E2_.ess$Ile13C ~ E2_.ess$Percent_Protein, data = E2_.ess) #determines whether there is a difference in d13CIle in E2 among Percent Protein diets
Ile
#follow it up by pairwise comparison to see which groups are diff. 
pairwise.wilcox.test(E2_.ess$Ile13C, E2_.ess$Percent_Protein,
                     p.adjust.method = "BH")

#Kruskal-Wallis ANOVA+Wilcox Rank Sum test
Leu <- kruskal.test(E2_.ess$Leu13C ~ E2_.ess$Percent_Protein, data = E2_.ess) #determines whether there is a difference in d13CLeu in E2 among Percent Protein diets
Leu
#follow it up by pairwise comparison to see which groups are diff. 
pairwise.wilcox.test(E2_.ess$Leu13C, E2_.ess$Percent_Protein,
                     p.adjust.method = "BH")

#Kruskal-Wallis ANOVA+Wilcox Rank Sum test
Val <- kruskal.test(E2_.ess$Val13C ~ E2_.ess$Percent_Protein, data = E2_.ess) #determines whether there is a difference in d13CVal in E2 among Percent Protein diets
Val
#follow it up by pairwise comparison to see which groups are diff. 
pairwise.wilcox.test(E2_.ess$Val13C, E2_.ess$Percent_Protein,
                     p.adjust.method = "BH")

#Kruskal-Wallis ANOVA+Wilcox Rank Sum test
Phe <- kruskal.test(E2_.ess$Phe13C ~ E2_.ess$Percent_Protein, data = E2_.ess) #determines whether there is a difference in d13CPhe in E2 among Percent Protein diets
Phe
#follow it up by pairwise comparison to see which groups are diff. 
pairwise.wilcox.test(E2_.ess$Phe13C, E2_.ess$Percent_Protein,
                     p.adjust.method = "BH")

#Kruskal-Wallis ANOVA+Wilcox Rank Sum test
Thr <- kruskal.test(E2_.ess$Thr13C ~ E2_.ess$Percent_Protein, data = E2_.ess) #determines whether there is a difference in d13CThr in E2 among Percent Protein diets
Thr
#follow it up by pairwise comparison to see which groups are diff. 
pairwise.wilcox.test(E2_.ess$Thr13C, E2_.ess$Percent_Protein,
                     p.adjust.method = "BH")

#Kruskal-Wallis ANOVA+Wilcox Rank Sum test
Lys <- kruskal.test(E2_.ess$Lys13C ~ E2_.ess$Percent_Protein, data = E2_.ess) #determines whether there is a difference in d13CLys in E2 among Percent Protein diets
Lys
#follow it up by pairwise comparison to see which groups are diff. 
pairwise.wilcox.test(E2_.ess$Lys13C, E2_.ess$Percent_Protein,
                     p.adjust.method = "BH")


```



```{r}

#Kruskal-Wallis ANOVA+Wilcox Rank Sum test
Ile <- kruskal.test(E2_.ess$Ile13C ~ E2_.ess$Type, data = E2_.ess) #determines whether there is a difference in d13CIle in E2 among Percent Protein diets
Ile
#follow it up by pairwise comparison to see which groups are diff. 
pairwise.wilcox.test(E2_.ess$Ile13C, E2_.ess$Type,
                     p.adjust.method = "BH")

#Kruskal-Wallis ANOVA+Wilcox Rank Sum test
Leu <- kruskal.test(E2_.ess$Leu13C ~ E2_.ess$Type, data = E2_.ess) #determines whether there is a difference in d13CLeu in E2 among Percent Protein diets
Leu
#follow it up by pairwise comparison to see which groups are diff. 
pairwise.wilcox.test(E2_.ess$Leu13C, E2_.ess$Type,
                     p.adjust.method = "BH")

#Kruskal-Wallis ANOVA+Wilcox Rank Sum test
Val <- kruskal.test(E2_.ess$Val13C ~ E2_.ess$Type, data = E2_.ess) #determines whether there is a difference in d13CVal in E2 among Percent Protein diets
Val
#follow it up by pairwise comparison to see which groups are diff. 
pairwise.wilcox.test(E2_.ess$Val13C, E2_.ess$Type,
                     p.adjust.method = "BH")

#Kruskal-Wallis ANOVA+Wilcox Rank Sum test
Phe <- kruskal.test(E2_.ess$Phe13C ~ E2_.ess$Type, data = E2_.ess) #determines whether there is a difference in d13CPhe in E2 among Percent Protein diets
Phe
#follow it up by pairwise comparison to see which groups are diff. 
pairwise.wilcox.test(E2_.ess$Phe13C, E2_.ess$Type,
                     p.adjust.method = "BH")

#Kruskal-Wallis ANOVA+Wilcox Rank Sum test
Thr <- kruskal.test(E2_.ess$Thr13C ~ E2_.ess$Type, data = E2_.ess) #determines whether there is a difference in d13CThr in E2 among Percent Protein diets
Thr
#follow it up by pairwise comparison to see which groups are diff. 
pairwise.wilcox.test(E2_.ess$Thr13C, E2_.ess$Type,
                     p.adjust.method = "BH")

#Kruskal-Wallis ANOVA+Wilcox Rank Sum test
Lys <- kruskal.test(E2_.ess$Lys13C ~ E2_.ess$Type, data = E2_.ess) #determines whether there is a difference in d13CLys in E2 among Percent Protein diets
Lys
#follow it up by pairwise comparison to see which groups are diff. 
pairwise.wilcox.test(E2_.ess$Lys13C, E2_.ess$Type,
                     p.adjust.method = "BH")

```

