---
title: "MixSIAR_d13C_Models"
author: "Conner Mertz"
date: "2023-10-31"
output: html_document
---

Script for running compound specific stable isotope mixing models of Amino Acid δ13C values from mice muscle (mixture/consumer) and dietary protein sources (casein for synthetic diet; cornmeal and cricket powder for semi-natural diet) to estimate microbial contributions of amino acids to host tissue.

The following code only evaluates mice fed the synthetic diet, but you simply repeat for mice in the semi-natural treatment and adjust the dietary protein sources accordingly. Lastly, this code can be used for all Essential Amino Acids (AA), but you must do one AA at a time.This code is for Ile, but you simply repeat for each respective AA. 


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Load MixSIAR package
```{r}
library(MixSIAR)
```

#Code used is from this tutorial:
https://github.com/brianstock/MixSIAR

#########Load mixture data
Load the mixture/consumer data, i.e. your:
- Consumer isotope values (d13C Ile muscle values from mice fed synthetic low, medium, and high protein diet treatments). 

```{r}

setwd("/Users/connermertz/Dropbox (vlab)/Conner Mertz/UNM biology/Chapter 1 E1 vs E2/Methods/SIA/E1_E2_d13C Data/Microbial_Contribution_MixingModel/Synthetic MIXSIAR/Ile")

# Load the mixture/consumer data
mix <- load_mix_data(filename="E1_muscle_consumer_Ile.csv", 
                     iso_names=c("d13CIle"), 
                     factors=c("Percent_Protein"), 
                     fac_random=c(TRUE), 
                     fac_nested=c(NULL), 
                     cont_effects=NULL) 
```


###Load source data
Load the source data, i.e. your:
- Source isotope values (dietary protein source; for this example its casein)

```{r}
setwd("/Users/connermertz/Dropbox (vlab)/Conner Mertz/UNM biology/Chapter 1 E1 vs E2/Methods/SIA/E1_E2_d13C Data/Microbial_Contribution_MixingModel/Synthetic MIXSIAR/Ile")

# Load the source data
source <- load_source_data(filename= "E1_muscle_sources_Ile.csv",
                           source_factors=NULL, 
                           conc_dep=FALSE, 
                           data_type="means", 
                           mix)
```

####Load discrimination data
See ?load_discr_data for details.

Note that Galloway et al. 2014 conducted feeding trials to create a “resource library”. In the mixing model, the sources are actually consumers fed exclusively each of the sources. This allowed them to set the discrimination = 0 (see isopod_discrimination.csv). Here we do the same. We assumed direct routing of AAESS and applied trophic discrimination factors (TDFs) of zero as done previously, Manlick & Newsome 2022.  

```{r}
setwd("/Users/connermertz/Dropbox (vlab)/Conner Mertz/UNM biology/Chapter 1 E1 vs E2/Methods/SIA/E1_E2_d13C Data/Microbial_Contribution_MixingModel/Synthetic MIXSIAR/Ile")

discr <- load_discr_data(filename="E1_muscle_discrimination_Ile.csv", mix)
```


##Plot data
This is your chance to check:

Are the data loaded correctly?
Is your mixture data in the source polygon?
Are one or more of your sources confounded/hidden?
When there are more than 2 biotracers, MixSIAR currently plots every pairwise combination. Here, that means (82)=28
 plots are produced. In the future, MixSIAR will offer non-metric multidimensional scaling (NMDS) plots for these cases.
 
 
```{r}
# Make an isospace plot
plot_data(filename="isospace_plot", plot_save_pdf=TRUE, plot_save_png=FALSE, mix,source,discr)
```
 
###Plot prior
Define your prior, and then plot using “plot_prior”

RED = your prior
DARK GREY = “uninformative”/generalist (alpha = 1)
LIGHT GREY = “uninformative” Jeffrey’s prior (alpha = 1/n.sources)

```{r}
# default "UNINFORMATIVE" / GENERALIST prior (alpha = 1)
plot_prior(alpha.prior=1,source)
```
##Write JAGS model file
In the Isopod Example we demo the “Residual only” error option. The differences between “Residual * Process”, “Residual only”, and “Process only” are explained in Stock and Semmens (2016). Also, looks like Alexi just did the residual as well, look into this further. 

```{r}
setwd("/Users/connermertz/Dropbox (vlab)/Conner Mertz/UNM biology/Chapter 1 E1 vs E2/Methods/SIA/E1_E2_d13C Data/Microbial_Contribution_MixingModel/Synthetic MIXSIAR/Ile")
# Write the JAGS model file
model_filename <- "MixSIAR_model.txt"
resid_err <- TRUE
process_err <- FALSE
write_JAGS_model(model_filename, resid_err, process_err, mix, source)
```

###Run model
First use run = "test" to check if 1) the data are loaded correctly and 2) the model is specified correctly:

```{r}
jags.1 <- run_model(run="test", mix, source, discr, model_filename)
```
##After a test run works, increase the MCMC run to a value that may converge
```{r}
jags.1 <- run_model(run="normal", mix, source, discr, model_filename)
```
jags.1 will be an rjags object where you can access the MCMC chains for plotting, aggregating sources a posteriori, etc.

####Analyze diagnostics and output
First you can set output options like file names, plot file types, etc. (see ?output_JAGS for details).

```{r}
output_options <- list(summary_save = TRUE,
                       summary_name = "summary_statistics",
                       sup_post = FALSE,
                       plot_post_save_pdf = TRUE,
                       plot_post_name = "posterior_density",
                       sup_pairs = FALSE,
                       plot_pairs_save_pdf = TRUE,
                       plot_pairs_name = "pairs_plot",
                       sup_xy = TRUE,
                       plot_xy_save_pdf = FALSE,
                       plot_xy_name = "xy_plot",
                       gelman = TRUE,
                       heidel = FALSE,
                       geweke = TRUE,
                       diag_save = TRUE,
                       diag_name = "diagnostics",
                       indiv_effect = FALSE,
                       plot_post_save_png = FALSE,
                       plot_pairs_save_png = FALSE,
                       plot_xy_save_png = FALSE,
                       diag_save_ggmcmc = FALSE)
```

Then you can call `output_JAGS` to process diagnostics, summary statistics, and create posterior density plots:

`output_JAGS(jags.1, mix, source, output_options)`

For a thorough explanation of the output from output_JAGS, see the Wolves Example section of the MixSIAR Manual. You will also find examples of accessing the MCMC chains for post hoc plotting and analysis there.

```{r}
output_JAGS(jags.1, mix, source, output_options)
```


